---
title: "R Notebook"
author: "Jerrick Tram"
date: "`r Sys.Date()`"
output: 
 html_notebook
---
Despite being more of a Disney/Pixar person myself, I did enjoy my fair share
of TV shows such as the Office, Game of Thrones, and Spongebob. Like most TV
shows where we had seasons we enjoyed, there were also seasons that just fell
flat. Are certain genres just not meant to have high ratings? Did the public
lose interest in shows that last too long? 

```{r message = FALSE}
library(ggthemes)   # Using the Economist theme
library(lubridate)  # Date-time manipulation
library(tidyverse)  # Data wrangling
```

```{r message = FALSE}
df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv")
head(df)
df$date <- as.Date(df$date)     # Convert to date
df <- df %>% 
  select(-titleId)              # Remove unnecessary column
```

Notice that shows have multiple genres. However, we'd like to know what
combination of genres tend to be popular.

```{r echo = TRUE, eval = TRUE}
df %>% 
  group_by(genres) %>% 
  summarize(n = n(),
            med_rating = median(av_rating),
            med_seasons = median((seasonNumber))) %>% 
  arrange(desc(n))
```

The people have spoken and it appears that shows that have "action-packed" or
"tense" themes are viewed the most. With solid ratings, the "lifespan" of these
shows are surprisingly short. Let's look at individual shows.

```{r echo = TRUE, eval = TRUE}
df2 <- df %>% 
  group_by(title, genres) %>% 
  summarize(seasons = max(seasonNumber),
            med_rating = median(av_rating),
            duration = as.numeric(max(date) - min(date))/365) %>% 
  arrange(desc(seasons))
```

44 seasons? Now that's quite an accomplishment. However, the median ratings 
are more volatile especially ABC Afterschool Specials being our outlier. 
Do more seasons contribute to the public giving lower ratings?

```{r echo = TRUE, eval = TRUE}
ggplot(df2, aes(seasons, med_rating, alpha = duration)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Seasons",
       y = "Median Rating",
       title = "Drama TV Show Ratings",
       alpha = "Duration") +
  theme_economist()
```

This doesn't tell us much. In fact, if we were going to run a regression...

```{r echo = TRUE, eval = TRUE}
tv_regression <- lm(med_rating ~ seasons, data = df2)
summary(tv_regression)
```

The number of seasons doesn't play a significant role in how the median ratings
turn out (with the exception of Spongebob as the quality clearly declined after
Season 3). So let's look exclusively at the top 15 shows if they stood the test
of time.

```{r message = FALSE}
top_15 <- df %>% 
  count(title, sort = TRUE) %>% 
  top_n(15)

top_15_stats <- inner_join(top_15, df, by = "title")

ggplot(top_15_stats, aes(date, av_rating, col = seasonNumber)) +
  geom_line(linetype = "solid") +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ title) +
  labs(x = "Date",
       y = "Average Rating",
       col = "Season Number",
       title = "The Top 15 Most Popular Drama TV Shows and Ratings Over Time") +
  theme_economist()
```