---
title: "Week 38 TidyTuesday: Cetaceans Data"
author: "Jerrick Tram"
date: "`r Sys.Date()`"
output: 
 html_notebook:
  toc: true
  toc_float: true
  theme: united
---
# Backstory
When I was a young kid, the first thing that comes up to my head thinking about
marine animals is Shamu, SeaWorld's most popular killer whale(s). You may have 
also heard of the movie *Free Willy* in the 90s going from a friendly picture
to the not so friendly parts of animal captivity. One thing I'd like to look 
at is the lifespan of marine mammals before and after 1972 due to The Marine 
Mammal Protection Act (MMPA) which regulated the location, time, and reasoning 
for capturing cetaceans.


# Cleaning
```{r echo = TRUE, eval = FALSE}
library(tidyverse)  # Data wrangling
library(lubridate)  # Datetime conversions

Cetaceans <- read_csv("allcetaceans.csv")
 
lapply(Cetaceans, class)   # Look for columns that need to be converted

Cetaceans <- Cetaceans %>% 
  select(-id, -notes, -transfer, -transfers, -transferDate, -name, -mother, -father) %>% 
  mutate_at(vars(originDate, entryDate, statusDate), as.Date, format = "%m/%d/%Y") 
  
Cetaceans$birthYear <- year(as.Date(Cetaceans$birthYear, format = "%Y"))


lapply(Cetaceans, class)   # Recheck columns
```

# Wrangling

```{r echo = TRUE, eval = TRUE}
Cetaceans <- Cetaceans %>% 
  mutate(lifespan = year(statusDate) - year(originDate)) %>% 
  mutate(mmpa = if_else(year(originDate) >= 1972, "Yes", "No"))
  

Cetaceans %>% 
  group_by(species, mmpa) %>% 
  count(mmpa, sort = TRUE) 
```
  
Roughly 10.4\% of bottlenose dolphins were acquired before the MMPA went into
effect. Since we talked about killer whales previously, they weren't so lucky
as 30.4\% of killer whales were acquired prior to the MMPA. So how does the
MMPA affect the lifespan of cetaceans?

```{r echo = TRUE, eval = TRUE}
Cetaceans %>% 
  group_by(species, acquisition, mmpa) %>% 
  summarize(avg_lifespan = mean(lifespan, na.rm = TRUE))
```

Something funky is going on as bottlenose dolphins have much shorter lifespans 
on average **AFTER** the MMPA went into effect. We need to investigate how 
they died whether it was through natural causes or outside environmental
factors.

```{r echo = TRUE, eval = TRUE}
Cetaceans %>% 
  filter(species == "Bottlenose") %>% 
  count(COD, sort = TRUE) 
```

Unfortunately, over a third of the bottlenose dolphins' cause of death (COD) is
unknown although some of the COD appears to be weather/suffering based. Does
location play a role?

```{r echo = TRUE, eval = TRUE}
Cetaceans %>% 
  filter(species == "Bottlenose") %>%
  group_by(mmpa) %>% 
  count(currently, acquisition, sort = TRUE)
```

Bottlenose dolphins are warm blooded creatures so ideally, they prefer cold
regions as with the majority of the locations based in Florida, Texas, and 
California, it's no wonder they didn't last long as those states can get
really hot.

# Visualization
```{r echo = TRUE, eval = TRUE}
Cetaceans %>% 
  select(species, acquisition, lifespan, mmpa, status) %>% 
  filter(species == "Bottlenose") %>% 
  group_by(acquisition, mmpa, status) %>% 
  summarize(mean_lifespan = mean(lifespan, na.rm = TRUE)) %>% 
  ggplot(aes(reorder(acquisition, mean_lifespan), mean_lifespan, fill = mmpa)) +
    geom_col(position = "dodge") +
    facet_wrap(~ status) +
    coord_flip() +
    labs(x = "Acquisition",
         y = "Average Lifespan",
         fill = "MMPA",
         title = "The Lifespan of Bottleneck Dolphins Before/After The MMPA")
```
