---
title: "Week 29 TidyTuesday: College Major and Income"
author: "Jerrick Tram"
date: "`r Sys.Date()`"
output: 
 html_notebook:
  toc: true
  toc_float: true
  theme: cerulean
---
Picking a major isn't easy for most students when considering job prospects,
income, work environment, etc. Let's not forget the cost of higher education
as those who unfortunately don't have scholarships or other funding sources
borrow approximately 5 to 6 figure loans to cover college expenses. As an
Economics major, I've been fortunate to live in Georgia as well as attending
an extremely affordable 4-year college with access to high quality professors.
Enough about me though, let's get to some exploring.

# Cleaning

```{r echo = TRUE, eval = TRUE}
library(tidyverse)

df <- read_csv("college.csv")
df <- df %>% 
  select(-Major_code, -Rank)   # Remove unnecessary columns
df
```

From an old [538 article](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/)
discussing the lucrativeness of certain college majors, STEM (dominated by 
Engineering) majors appear frequently as the top money makers. However, 
does money affect how many people choose said major?

# EDA

```{r echo = FALSE, eval = TRUE}
df %>% 
  group_by(Category) %>% 
  summarize(total = sum(Total, na.rm = TRUE),
            median_salary = median(Median)) %>% 
  arrange(desc(total)) 
```

An overwhelming number of students picked business yet the rest of the top
5/10 majors falls under the Humanities and Liberal Arts. The data was collected
in 2010-2010 although it's interesting to see that money isn't the sole
motivator towards picking a major. However, when reality kicks in, we also need
to look at employability and the one thing college students fear most isn't
frictional unemployment but **underemployment.**

```{r echo = TRUE, eval = TRUE}
ue <- df %>% 
  group_by(Major, ShareWomen, Category) %>% 
  filter(!is.na(ShareWomen)) %>% 
  summarize(total = sum(Employed, na.rm = TRUE),
            underemployed = sum(Low_wage_jobs, na.rm = TRUE),
            underemployed_percentage = underemployed/total) %>% 
  arrange(desc(underemployed_percentage)) 
```

Turns out the stereotypes of arts struggling in the job market aren't exactly
false. A whopping 36.5\% of students that majored in Cosmetology Services and
Culinary Arts are underemployed. Economics isn't doing too bad for itself at 
10.2\%. As you go further down the list, you start to see majors that are more
specialized with unique niches (or STEM) but less representation from women. 
Do we have fewer than 50\% of women in STEM?

```{r echo = TRUE, eval = TRUE}
df %>% 
  group_by(Category) %>% 
  summarize(women_representation = median(ShareWomen, na.rm = TRUE))
```

The sciences are where there's decent representation of women but as for
engineering, computers and mathematics, it just shows 26.9\% and 22.7\%
respectively. Now let's make some data visualizations

# Visualization

```{r echo = TRUE, eval = TRUE}
library(ggthemes)

df %>% 
  ggplot(aes(fct_reorder(Category, Median), Median, na.rm = TRUE)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = "Discipline",
       y = "Median Salary",
       title = "2010 - 2012 Median Salaries by Discipline") +
  theme_economist()
```

```{r echo = TRUE, eval = TRUE}
ue %>% 
  ggplot(aes(underemployed_percentage, ShareWomen, size = underemployed)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ Category) +
  labs(x = "Underemployment Percentage",
       y = "Percentage of Women",
       size = "Total Underemployed",
       title = "Underemployment by Discipline") +
  theme_economist()
```